Pretty darn good control: when are approximate solutions better than approximate models

### 2023-01-17 manuscript/manuscript.Rmd
3d model
bit more 3d model

Most commonly, fisheries use 1-dimensional models, $n=1$.
One such model appearing in a variety of ecological contexts is *logistic growth*, for which
$$
  f(N_t) = r N_t\big(1 - N_t / K \big) =: L(N_t;\ r, K).
$$
More relevant to our current work, is the following single-species model studied in \cite{may77},
\begin{align}
  \label{eq:may}
  N_{t+1} - N_t
  = L(N_t) - F(N_t, H; \beta,c),
\end{align}
where,
\begin{align*}
  F(N,H;\ \beta,c) := \frac{\beta H N^2}{c^2 + N^2}.
\end{align*}
The model has five parameters: a constant population $H$ of a species which preys on $N$, the maximal rate of predation $\beta$, and the predation half-maximum biomass $c$.

@@

Consider the following three species *generalization* of eq. \eqref{eq:may}:
\begin{align}
  \label{eq:may 3d}
  \Delta X_t = L(X_t; r_X, K_X)
\end{align}

>>>>

Most commonly, fisheries use 1-dimensional models, $n=1$.
One such model appearing in a variety of ecological contexts is *logistic growth*, for which
$$
  f(X_t) = r X_t\big(1 - X_t / K \big) =: L(X_t;\ r, K).
$$
More relevant to our current work, is the following single-species model studied in \cite{may77},
\begin{align}
  \label{eq:may}
  X_{t+1} - X_t
  = L(X_t;\ r, K) - F(X_t, H; \beta,c),
\end{align}
where,
\begin{align*}
  F(X,H;\ \beta,c) := \frac{\beta H X^2}{c^2 + X^2}.
\end{align*}
The model has five parameters: the growth rate $r$ and carrying capacity $K$ for $X$, a constant population $H$ of a species which preys on $X$, the maximal rate of predation $\beta$, and the predation half-maximum biomass $c$.

@@

Consider the following three species *generalization* of eq. \eqref{eq:may}:
\begin{align}
\label{eq:3d model}
\begin{split}
  \Delta X_t &= L(X_t;\ r_X, K_X) - F(X_t, Z_t;\ \beta, c) - c_{XY}X_tY_t - h_t,\\
  \Delta Y_t &= L(Y_t;\ r_Y, K_Y) - D F(Y_t, Z_t;\ \beta, c) - c_{XY}X_tY_t,\\
  \Delta Z_t &= \left(f(X_t + DY_t\right) - d_Z)Z_t.
\end{split}
\end{align}
The model is conceptualized in Fig. \ref{fig:tbd}.
It contains three populations, the state at time $t$ is $(X_t, Y_t, Z_t)$. 
Species $Z$ preys on both $X$ and $Y$, while the latter two compete for resources.
There are ten parameters in this model: 
The growth rate and carrying capacity, $r_X$, $K_X$, $r_Y$ and $K_Y$, of $X$ and $Y$.
A parameter $c_{XY}$ mediating a Lotka-Volterra competition between $X$ and $Y$.
A maximum hunting intensity $\beta$ and half-maximum $c$ specifying how $Z$ forages on $X$ and $Y$.
A parameter $D$ regulating a relative preference of $Z$ to prey on $Y$.
Finally, a death rate $d_Z$ and a parameter $f$ related to the birth rate of $Z$.

This model generalizes eq. \eqref{eq:may} in the following sense:
If $f=d_Z=c_{XY}=0$, and we let $Z_{t=0}=H$, then eq. \eqref{eq:3d model} becomes
\begin{align*}
  \Delta X_t &= L(X_t;\ r_X, K_X) - F(X_t, H;\ \beta, c) - h_t,\\
  \Delta Y_t &= L(Y_t;\ r_Y, K_Y) - D F(Y_t, H;\ \beta, c),\\
  Z_t &= H.
\end{align*}
The equations for $X_t$ and $Y_t$ thus become decoupled, with $X_t$ evolving according to \eqref{eq:may}.
